---
title: "Read and Analyze earthquakes"
author: "Alan Jackson"
format: html
editor: source
---

##  Read in and analyze earthquake data

Read in

Trim to diocese

Find nearest parishes to each epicenter

```{r setup}

library(tidyverse)

path <- "/home/ajackson/Dropbox/Rprojects/TexasDioceseCreationCare/Data/"
out_path <- "/home/ajackson/Dropbox/projects/CeationCareTaskforce/Maps_and_tables/Earthquake/"

googlecrs <- "EPSG:4326"

Convocations <- readRDS(paste0(path,"Convocations.rds"))  

Parishes <- readRDS(paste0(path, "Parishes.rds")) %>% 
  mutate(City=stringr::str_trim(City) ) %>% 
  mutate(Label=stringr::str_trim(Label) )

```

### Read in data and trim

```{r}

Quakes <- read_csv(paste0(path, "Earthquake/texnet_events.csv")) %>% 
  select(Date='Origin Date',
         Magnitude='Local Magnitude',
         Depth='Depth of Hypocenter (Km. Rel to Ground Surface)',
         Lat='Latitude (WGS84)',
         Lon='Longitude (WGS84)') %>% 
         sf::st_as_sf(coords=c("Lon", "Lat"), crs=googlecrs)

#   Combine the polygons into one

Diocese <- Convocations %>% 
  mutate(Label="None") %>% 
  select(-County) %>% 
  sf::st_union() %>% 
  sf::st_make_valid()

plot(Diocese)

a <- sf::st_intersects(Convocations, Quakes, sparse = TRUE)
a <- unlist(replace(a, !sapply(a, length), 209))
a <- a[a<209]

Quakes_2 <- Quakes[a, ]


```

###       Nearest churches

The intensity decreases as the square of the distance from the quake at depth,
so if we say that dropping the intensity by one order of magnitude (so a mag 4
becomes like a mag 3) then the distance on the earth will be 3 times the depth. 

Or better, use the depth and the reduction by the square of the distance law 
determine how far from the epicenter intensity drops to mag 3.

$radius = depth*\sqrt{10^{(mag-3)}-1}$


```{r}

Quakes_2 <- Quakes_2 %>% 
  mutate(Radius=1000*Depth*sqrt(10**(Magnitude-3)-1)) %>% 
  select(Date, Magnitude, Depth, Radius, geometry) %>% 
  sf::st_as_sf()

#   Define some Radius polygons around each epicenter
# Quake_polys <- Quakes_2 %>% 
#   sf::st_buffer(Radius) %>% 
#   sf::st_as_sf() # Units are meters

Quake_polys <- sf::st_buffer(Quakes_2, Quakes_2$Radius)

tmap::tmap_options(basemaps="OpenStreetMap")

tmap::tmap_mode("view") # set mode to interactive plots

# map <- 
  tmap::tm_shape(Quake_polys) + 
  tmap::tm_fill(col = "red", title = "Quake influences", alpha=0.5, style="pretty")+
  tmap::tm_borders(lwd=0.2) +
  tmap::tm_shape(Parishes) + 
  tmap::tm_dots()

#   Make map showing effect radius and magnitude

bbox <- sf::st_bbox(Quake_polys)

Base_basemapR <- basemapR::base_map(bbox, basemap="mapnik", increase_zoom=2)

Quake_polys %>% 
  ggplot(aes(fill="red", alpha=0.3)) +
  Base_basemapR +
  geom_sf(color="black", shape=1 ) +
  geom_sf(alpha=0.5, shape=21 ) +
  geom_sf(data=Parishes) +
  # scale_color_gradientn(colors=RColorBrewer::brewer.pal(9, "YlOrRd"),
  #                        name="Fire Risk",
  #                        na.value = "grey100") +
  labs(title=paste("Historical earthquakes and affected areas"),
       x="Longitude",
       y="Latitude") +
  theme(legend.position="none") +
  annotate("text", label=
           "Data sourced from Texnet Earthquake Catalog
           Texas Bureau of Economic Geology
           Map a product of Alan Jackson", 
           x=-Inf, 
           y=-Inf,
           hjust="inward", vjust="inward", size = 2) +
  coord_sf(xlim=c(bbox$xmin, bbox$xmax),c(bbox$ymin, bbox$ymax))

  ggsave(paste0(out_path, "Earthquakes.jpg"))

```




