---
title: "First Street Flood"
author: "Alan Jackson"
format: html
editor: visual
---

## Read in and analyze the First Street Foundation flood data

Data acquired from FSF through a request on an AWS server After access granted, copied to a bucket on my AWS account Then downloaded (one at a time) the files to my PC

```{r setup}

library(tidyverse)

path <- "/home/ajackson/Dropbox/Rprojects/TexasDioceseCreationCare/Data/"
output_path <- "/home/ajackson/Dropbox/projects/CeationCareTaskforce/FirstStreetFlood/"

#   Some useful data

Census <- readRDS(paste0(path, "Census_data_for_Texas_by_tract.rds"))

Convocations <- readRDS(paste0(path, "Convocation_and_tracts.rds"))
Parishes <- readRDS(paste0(path, "Parishes_plus_Census_tracts.rds"))

```

## Make little graphs of flood factor for plot insets

```{r factor insets}

FF3 <- tribble(~Depth, ~Risk,
               12.,     0.01,
               7.5,     0.05,
               4.5,     0.08,
               2.0,     0.12)

FF5 <- tribble(~Depth, ~Risk,
               30.0,   0.03,
               18.0,   0.09,
               8.0,    0.20,
               5.0,    0.30,
               2.0,    0.40 )

FF8 <- tribble(~Depth, ~Risk,
               30.0,   0.30,
               18.0,   0.40,
               10.0,   0.70,
               4.0,    0.98
              )

pFF3 <- FF3 %>% ggplot(aes(x=Risk, y=Depth)) +
  geom_point(size=0.5) +
  geom_smooth(se=FALSE, span=1, size=0.5) +
  theme(plot.title = element_text(size = 4),
        axis.text.x = element_text(size = 4),
        axis.text.y = element_text(size = 4),
        axis.title.x = element_text(size = 4),
        axis.title.y = element_text(size = 4))+
  labs(title="Flood Factor 3",
       x="30-Year Cumulative Risk",
       y="Flood Depth (in)") +
  scale_x_continuous(labels = scales::percent)

pFF5 <- FF5 %>% ggplot(aes(x=Risk, y=Depth)) +
  geom_point(size=0.5) +
  geom_smooth(se=FALSE, span=1, size=0.5) +
  theme(plot.title = element_text(size = 4),
        axis.text.x = element_text(size = 4),
        axis.text.y = element_text(size = 4),
        axis.title.x = element_text(size = 4),
        axis.title.y = element_text(size = 4))+
  labs(title="Flood Factor 5",
       x="30-Year Cumulative Risk",
       y="Flood Depth (in)") +
  scale_x_continuous(labels = scales::percent)

pFF8 <- FF8 %>% ggplot(aes(x=Risk, y=Depth)) +
  geom_point(size=0.5) +
  geom_smooth(se=FALSE, span=1, size=0.5) +
  theme(plot.title = element_text(size = 4),
        axis.text.x = element_text(size = 4),
        axis.text.y = element_text(size = 4),
        axis.title.x = element_text(size = 4),
        axis.title.y = element_text(size = 4))+
  labs(title="Flood Factor 8",
       x="30-Year Cumulative Risk",
       y="Flood Depth (in)") +
  scale_x_continuous(labels = scales::percent)

```


## Read the data in by census tract and filter out diocese

```{r}

df <- read_csv(paste0(path, "FSF_Flood/fsf_flood_tract_summary.csv")) %>% 
  rename(GEOID=fips)

#   Add polygons and census data

df2 <- inner_join(df, Census, by="GEOID") %>%
  sf::st_as_sf() %>% 
  sf::st_make_valid()

#   Add Convocations and filter out data ouside the diocese

Convocations <- Convocations %>% 
  sf::st_drop_geometry() %>% 
  unnest_longer(Tract) %>% 
  filter(Tract != "48071710600") %>% # unoccupied spoil banks in Galveston Bay
  rename(GEOID=Tract) 

Convocations <- left_join(Convocations, df2, by="GEOID")

# Tract_geoms <- Census %>% 
#   select(Tract=GEOID)  

Convocations <- Convocations %>% 
  mutate(Name=stringr::str_remove(Name, " Convocation| Region")) %>% 
  # left_join(., Tract_geoms, by="Tract") %>% 
  sf::st_as_sf()

# df2 %>%
#   mutate(Fact5=count_floodfactor5/count_property*100) %>% 
#   # mutate(count_floodfactor10=pmin(count_floodfactor10, 100)) %>% 
#   ggplot() + 
#   geom_histogram(aes(x=Fact5))

df3 <- Convocations %>% 
  mutate(Fact5=count_floodfactor5/count_property*100)  
  

#   make a diagnostic map

tmap::tmap_options(basemaps="OpenStreetMap")

tmap::tmap_mode("view") # set mode to interactive plots

tmap::tm_shape(df3) + 
    #tmap::tm_sf(col="ConvoColor", alpha=0.3) +
  tmap::tm_fill(col = "Fact5", title = "Flood Factor 5 Percent", alpha=0.3, style="pretty")+
  tmap::tm_borders(lwd=0.1) 



```

## Make maps of flood factor 3, 5, and 8 colored by percent houses affected for the diocese

```{r whole diocese}

foo <- Convocations %>% 
  arrange(GEOID) %>% 
  filter(!duplicated(GEOID)) %>% 
  mutate(Fact3=count_floodfactor3/count_property*100) %>%  
  mutate(Fact5=count_floodfactor5/count_property*100) %>% 
  mutate(Fact8=count_floodfactor8/count_property*100)  

foo %>% ggplot(aes(x=Fact3)) +
  geom_histogram()

foo <- foo %>% 
  mutate(Fact3_grp = cut(Fact3,
                     breaks=c(0,5,10,20,40,100),
                     labels=c("0%-5%", "5%-10%", "10%-20%", "20%-40%", "40%-100%"),
                     include.lowest = TRUE)) %>% 
  mutate(Fact5_grp = cut(Fact5,
                     breaks=c(0,5,10,20,40,100),
                     labels=c("0%-5%", "5%-10%", "10%-20%", "20%-40%", "40%-100%"),
                     include.lowest = TRUE)) %>% 
  mutate(Fact8_grp = cut(Fact8,
                     breaks=c(0,5,10,20,40,100),
                     labels=c("0%-5%", "5%-10%", "10%-20%", "20%-40%", "40%-100%"),
                     include.lowest = TRUE))

  box <- foo %>%
    sf::st_bbox() 
  
  Base_basemapR <- basemapR::base_map(box, basemap="mapnik", increase_zoom=2)
  
 #Colour Palette
pal <- c(
  "40%-100%" = "red",
  "20%-40%" = "orange", 
  "10%-20%" = "yellow", 
  "5%-10%" = "lightgreen",
  "0%-5%" = "white" 
)                          

for (map_var in c("Fact3_grp", "Fact5_grp", "Fact8_grp")){
  
  if (map_var == "Fact3_grp") {Plot_inset <- pFF3}
  if (map_var == "Fact5_grp") {Plot_inset <- pFF5}
  if (map_var == "Fact8_grp") {Plot_inset <- pFF8}
  
  
  Plot_title <- case_when(
    map_var == "Fact3_grp" ~ "Flood Factor 3 by Census Tract",
    map_var == "Fact5_grp" ~ "Flood Factor 5 by Census Tract",
    map_var == "Fact8_grp" ~ "Flood Factor 8 by Census Tract"
  )
  map_var <- ensym(map_var)
  
     p <- 
      foo %>% 
      ggplot() +
      Base_basemapR +
      geom_sf(aes(fill=!!map_var) ,alpha=0.6, lwd=0.0) +
      scale_fill_manual( values=pal,
                         limits=names(pal),
                         name="Homes flooded") +
      annotate("text", label="
             Data sourced from First Street flood Model, June 2020 
             Map a product of Alan Jackson", 
             x=-Inf, 
             y=-Inf,
             hjust="inward", vjust="inward", size = 1.5)+
      labs(title=Plot_title) +
      annotation_custom(
            grob = ggplotGrob(Plot_inset),
            xmin = -96.5, xmax = -93.8, ymin = 33, ymax = 34.5
    ) 
    
      ggsave(paste0(output_path, "FirstStreet_Diocese_",
                    paste(Plot_title, sep="_"),
                    ".jpg"))
     # print(p)
}
  
```


## Make maps of flood factor 3, 5, and 8 colored by percent houses affected for the convocations

```{r convocations}

Inset_position <- tribble(~Convo, ~xmin, ~xmax, ~ymin, ~ymax,
                          "Austin",  -97.3, -96.3, 30.9, 31.7,
                          "Central", -96.2, -95.5, 31.0, 31.6,
                          "East Harris", -95.0, -94.7, 29.9, 30.1,
                          "Galveston",   -94.7, -94.4, 29.5, 29.8,
                          "Northeast",   -94.3, -93.5, 30.7, 31.25,
                          "North",       -96.8, -95.6, 33.5, 34.5,
                          "Northwest",   -98.5, -97.75, 31.4, 32.0,
                          "San Jacinto", -94.9, -94.2, 30.6, 31.2,
                          "Southeast",   -95.1, -94.5, 30.75, 31.25,
                          "Southwest",   -96.75, -96.25, 28.9, 29.4,
                          "West Harris", -96.1, -95.85, 29.85, 30.05)

foo <- Convocations %>% 
  mutate(Fact3=count_floodfactor3/count_property*100) %>%  
  mutate(Fact5=count_floodfactor5/count_property*100) %>% 
  mutate(Fact8=count_floodfactor8/count_property*100)  

foo <- foo %>% 
  mutate(Fact3_grp = cut(Fact3,
                     breaks=c(0,5,10,20,40,100),
                     labels=c("0%-5%", "5%-10%", "10%-20%", "20%-40%", "40%-100%"),
                     include.lowest = TRUE)) %>% 
  mutate(Fact5_grp = cut(Fact5,
                     breaks=c(0,5,10,20,40,100),
                     labels=c("0%-5%", "5%-10%", "10%-20%", "20%-40%", "40%-100%"),
                     include.lowest = TRUE)) %>% 
  mutate(Fact8_grp = cut(Fact8,
                     breaks=c(0,5,10,20,40,100),
                     labels=c("0%-5%", "5%-10%", "10%-20%", "20%-40%", "40%-100%"),
                     include.lowest = TRUE))

 #Colour Palette
pal <- c(
  "40%-100%" = "red",
  "20%-40%" = "orange", 
  "10%-20%" = "yellow", 
  "5%-10%" = "lightgreen",
  "0%-5%" = "white" 
)                     

Convo_names <- Convocations$Label %>% unique()

  box <- foo %>%
    sf::st_bbox() 
  
  Base_basemapR <- basemapR::base_map(box, basemap="mapnik", increase_zoom=2)
  
for (Convo in Convo_names) {
  
  Inset <- Inset_position[Inset_position$Convo==Convo,]
  
  foobar <- foo %>% 
    filter(Label == Convo)
  
  box <- foobar %>%
    sf::st_bbox() 
  
  Base_basemapR <- basemapR::base_map(box, basemap="mapnik", increase_zoom=2)
  
  Convo_parishes <- Parishes %>% 
    mutate(Convocation=stringr::str_remove(Convocation, " Convocation| Region")) %>% 
    filter(stringr::str_equal(Convocation, Convo))
  
  for (map_var in c("Fact3_grp", "Fact5_grp", "Fact8_grp")){
    
    if (map_var == "Fact3_grp") {Plot_inset <- pFF3}
    if (map_var == "Fact5_grp") {Plot_inset <- pFF5}
    if (map_var == "Fact8_grp") {Plot_inset <- pFF8}
  
    Plot_label <- case_when(
      map_var == "Fact3_grp" ~ "10% chance\nof 6\" flood",
      map_var == "Fact5_grp" ~ "10% chance\nof 20\" flood",
      map_var == "Fact8_grp" ~ "35% chance\nof 20\" flood"
    )
    Plot_title <- case_when(
      map_var == "Fact3_grp" ~ "Flood Factor 3 by Census Tract",
      map_var == "Fact5_grp" ~ "Flood Factor 5 by Census Tract",
      map_var == "Fact8_grp" ~ "Flood Factor 8 by Census Tract"
    )
    map_var <- ensym(map_var)
       
    # p <- 
        foobar %>% 
        ggplot() +
        Base_basemapR +
        annotation_custom(
              grob = ggplotGrob(Plot_inset),
              xmin = Inset$xmin, xmax = Inset$xmax, ymin = Inset$ymin, ymax = Inset$ymax
      ) +
        geom_sf(aes(fill=!!map_var) ,alpha=0.6, lwd=0.0) +
        scale_fill_manual( values=pal,
                           limits=names(pal),
                           name="Homes flooded") +
        geom_sf(data=Convo_parishes, show.legend=FALSE) +
        ggsflabel::geom_sf_label_repel(data=Convo_parishes, aes(label=Label), 
                                       size=2, max.overlaps=20, force=80) +
        annotate("text", label="
               Data sourced from First Street flood Model, June 2020 
               Map a product of Alan Jackson", 
               x=-Inf, 
               y=-Inf,
               hjust="inward", vjust="inward", size = 1.5)+
        labs(title=Plot_title,
             subtitle=Convo) 
      
        ggsave(paste0(output_path, "FirstStreet_Convo_",
                      paste(Plot_title, Convo, sep="_"),
                      ".jpg"))
        # print(p)
  }
}
  
```








