---
title: "Read Storm Surge data"
author: "Alan Jackson"
format: html
editor: visual
---

## Setup

Data downloaded from <https://www.nhc.noaa.gov/nationalsurge/>

Zachry, B. C., W. J. Booth, J. R. Rhome, and T. M. Sharon, 2015: A National View of Storm Surge Risk and Inundation. Weather, Climate, and Society, 7(2), 109--117. DOI: http://dx.doi.org/10.1175/WCAS--D--14--00049.1

-   Data comes in NAD83

Each dataset contains an ESRI World File (.tfw) and metadata .xml file. These GeoTIFFs are 8-bit unsigned integer raster datasets that correspond to 1 ft inundation bins (e.g., Class Value 1 corresponds to the 0-1 ft inundation bin, Class Value 2 corresponds to the 1-2 ft inundation bin, and so on). The maximum Class Value is 21, and inundation in excess of 20 ft is assigned a Class Value of 21. A Class Value of 99 is assigned to leveed areas. A more detailed description of the data can be found in the associated metadata.

```{r setup}

library(tidyverse)
library(stars)
library(terra)
library(tidyterra)

googlecrs <- "EPSG:4326"
localUTM <- "EPSG:32615"
inputcrs <- "EPSG:4269" # from metadata file

path <- "/home/ajackson/Dropbox/Rprojects/TexasDioceseCreationCare/Data/"

Parishes <- readRDS(paste0(path, "Parishes.rds"))

#   Define an AOI
bbox <- sf::st_bbox(sf::st_transform(Parishes, crs=inputcrs))

bbox["xmin"] <- -96.38
bbox["xmax"] <- -93.66
bbox["ymin"] <- 28.39
bbox["ymax"] <- 30.47

#   Expand box by 20% to give a little extra room
expand_box <- function(bbox, pct=0.2){
  Dx <- (bbox[["xmax"]]-bbox[["xmin"]])*pct
  Dy <- (bbox[["ymax"]]-bbox[["ymin"]])*pct
  bbox["xmin"] <- bbox["xmin"] - Dx
  bbox["xmax"] <- bbox["xmax"] + Dx
  bbox["ymin"] <- bbox["ymin"] - Dy
  bbox["ymax"] <- bbox["ymax"] + Dy
  return(bbox)
}

```

## Read in the files and clip to AOI

```{r read in and clip}

# tif = system.file(paste0(path, "US_SLOSH_MOM_Inundation_v3/us_Category1_MOM_Inundation_HIGH.tif"), package = "stars")
# x <- stars::read_stars(paste0(path, "US_SLOSH_MOM_Inundation_v3/us_Category1_MOM_Inundation_HIGH.tif")) 
x <- terra::rast(paste0(path, "US_SLOSH_MOM_Inundation_v3/us_Category1_MOM_Inundation_HIGH.tif")) 
# par(mar = rep(0,4))
# plot(x, col = grey((4:10)/10))

#   Clip to AOI

# cropped <-  sf::st_crop(x, bbox)
cropped <- terra::crop(x, bbox)
# par(mar = rep(0,4))
# plot(cropped, col = grey((4:10)/10))
# plot(cropped, col = sf::sf.colors(6))
# plot(sf::st_crop(sf::st_transform(Parishes, crs=inputcrs), bbox), col="black")

# pt1 <- sf::st_transform(Parishes, crs=inputcrs)[62,] %>% 
#   st_buffer(100) %>% st_as_sf() # 1000 equals a diameter of 1 mile, somehow.
Parishes <- Parishes %>% 
  mutate(ID=row_number())

pts <- sf::st_transform(Parishes, crs=inputcrs) %>% 
  st_buffer(100) %>% st_as_sf() # 1000 equals a diameter of 1 mile, somehow.

# pt <- sf::st_transform(Parishes, crs=inputcrs)[62,]
# 
# box <- sf::st_transform(Parishes, crs=inputcrs)[62,] %>% 
#   st_buffer(1200) %>% st_as_sf() %>% st_bbox()
# 
# Base_basemapR <- basemapR::base_map(box, basemap="mapnik", increase_zoom=2)

# pt1 %>% 
#   ggplot() +
#   Base_basemapR +
#   geom_sf(aes(alpha=0.1))

ggplot() +
  geom_spatraster(data=cropped) +
  geom_sf(data=sf::st_crop(sf::st_transform(Parishes, crs=inputcrs), bbox),
          color="black")

#   Extract values at church locations

foo <- cropped %>% 
  terra::extract(vect(pts), method="bilinear", ID=TRUE) %>% 
    group_by(ID) %>% 
    summarize(value=mean(data_range, na.rm=TRUE)) %>% 
  filter(!is.nan(value)) %>% 
  filter(value<99) %>% 
  inner_join(., Parishes, by="ID") %>% 
  st_as_sf()

box <- sf::st_transform(foo, crs=inputcrs) %>%
  st_as_sf() %>% st_bbox() %>% expand_box(., 0.5)

Base_basemapR <- basemapR::base_map(box, basemap="mapnik", increase_zoom=3)


st_as_sf(foo) %>% 
  ggplot() +
  Base_basemapR +
  geom_sf(color="black") +
  # geom_sf_label(aes(label=paste(stringr::str_remove(Name, "Episcopal Church"), signif(value,2)))) +
  ggsflabel::geom_sf_label_repel(aes(label=paste(stringr::str_remove(Name, "Episcopal Church"), signif(value,2), "ft"))) +
  labs(title="SLOSH model from NOAA, Cat 1 Hurricane Storm Surge in feet",
       x="Longitude",
       y="Latitude") +
  coord_sf(xlim=c(box$xmin, box$xmax),c(box$ymin, box$ymax))

ggsave("/home/ajackson/Desktop/Cat_1_slosh_model.jpg")

# plot.new()
# plot(terra::crop(cropped, pt1))
# points(vect(pt))

```

# Make contours (polygons) from the geotiff grid

```{r contour}

Lowres <- cropped %>% 
  aggregate(fact=10, fun="median", na.rm=TRUE) %>% 
  filter(data_range<99)

ggplot() +
  geom_spatraster(data=Lowres) +
  geom_sf(data=sf::st_crop(sf::st_transform(Parishes, crs=inputcrs), bbox),
          color="black")

Tinybox <- bbox 
Tinybox["xmin"] <- -94.868534
Tinybox["xmax"] <- -94.749265
Tinybox["ymin"] <- 29.239547
Tinybox["ymax"] <- 29.346297

ggplot() +
  basemapR::base_map(Tinybox, basemap="mapnik", increase_zoom=3)+
  geom_spatraster(data=terra::crop(Lowres, Tinybox), alpha=0.5)

# foo_sm <- terra::focal(foo_terra, w=3, fun=mean)

LowresSm <- Lowres %>% 
  terra::crop(Tinybox) %>% 
  terra::focal(w=5, fun=mean, na.rm=TRUE) 

Contours <- LowresSm %>% 
  mapiso::mapiso() %>% 
  smoothr::smooth(method="ksmooth", smoothness=2)
  # smoothr::smooth(mapiso::mapiso(terra::crop(Lowres, Tinybox)), 
  #                           method="ksmooth", smoothness=2)

ggplot() +
  basemapR::base_map(Tinybox, basemap="mapnik", increase_zoom=3)+
  geom_sf(data=Contours, color="black") +
  geom_spatraster(data=terra::crop(LowresSm, Tinybox), alpha=0.2)+
  scale_fill_gradientn(colors=rainbow(5), limits=c(0,10), na.value="transparent") +
  geom_sf(data=sf::st_crop(sf::st_transform(Parishes, crs=inputcrs), Tinybox),
          color="black")


```

##        Levee polygons

```{r levees}

Levees <- cropped %>% 
  aggregate(fact=10, fun="median", na.rm=TRUE) %>% 
  filter(data_range>98) %>% 
  as.polygons() %>% 
  st_as_sf() %>% 
  st_make_valid() %>% 
  st_cast("POLYGON") %>% 
  mutate(Levee_ID=row_number()) %>% 
  mutate(Levee_name=case_when(
    Levee_ID==1 ~ "Port Arthur",
    Levee_ID==2 ~ "Texas City",
    Levee_ID==3 ~ "Freeport",
    Levee_ID==4 ~ "Matagorda"
  )) %>% 
  select(Levee_ID, Levee_name, geometry) 

tmap::tmap_options(basemaps="OpenStreetMap")

tmap::tmap_mode("view") # set mode to interactive plots


tmap::tm_shape(Levees) + 
  tmap::tm_fill(title = "Levees", alpha=0.5, style="cat")+
  tmap::tm_text("Levee_name")+
  tmap::tm_borders()

saveRDS(Levees, paste0(path, "StormSurge_Levee_Polygons.rds"))
```

##        Which churches are behind levees?

```{r levee protected}

Protected <- st_join(pts, Levees, join=st_within) %>% 
  filter(!is.na(Levee_ID))

tmap::tmap_options(basemaps="OpenStreetMap")

tmap::tmap_mode("view") # set mode to interactive plots

tmap::tm_shape(Protected) + 
  tmap::tm_fill(title = "Levees", alpha=0.5, style="cat")+
  tmap::tm_text("Name")+
  tmap::tm_borders()


```











