---
title: "Sea Level Rise"
author: "Alan Jackson"
format: html
editor: visual
---

## NOAA sea level data

Data downloaded from https://coast.noaa.gov/slrdata/

Metadata:

These data were created as part of the National Oceanic and Atmospheric
Administration Office for Coastal Management's efforts to create an online
mapping viewer depicting potential sea level rise and its associated impacts on
the nation's coastal areas. The purpose of the mapping viewer is to provide
coastal managers and scientists with a preliminary look at sea level rise (slr)
and coastal flooding impacts. The viewer is a screening-level tool that uses
nationally consistent data sets and analyses. Data and maps provided can be used
at several scales to help gauge trends and prioritize actions for different
scenarios. The Sea Level Rise and Coastal Flooding Impacts Viewer may be
accessed at:

https://www.coast.noaa.gov/slr

These data depict the potential inundation of coastal areas resulting from
current Mean Higher High Water (MHHW) conditions. The process used to produce
the data can be described as a modified bathtub approach that attempts to
account for both local/regional tidal variability as well as hydrological
connectivity. The process uses two source datasets to derive the final
inundation rasters and polygons and accompanying low-lying polygons: the Digital
Elevation Model (DEM) of the area and a tidal surface model that represents
spatial tidal variability. The tidal model is created using the NOAA National
Geodetic Survey's VDATUM datum transformation software (http://vdatum.noaa.gov)
in conjunction with spatial interpolation/extrapolation methods and represents
the MHHW tidal datum in orthometric values (North American Vertical Datum of
1988).

The model used to produce these data does not account for erosion, subsidence,
or any future changes in an area's hydrodynamics. It is simply a method to
derive data in order to visualize the potential scale, not exact location, of
inundation from sea level rise.

The purpose of these data is to show potential inundation from current Mean
Higher High Water (MHHW) for the area.

A detailed methodology for producing these data can be found via the following url:
https://coast.noaa.gov/data/digitalcoast/pdf/slr-inundation-methods.pdf

These data illustrate the scale of potential flooding, not the exact location,
and do not account for erosion, subsidence, or future construction. Inundation
is shown as it would appear during the highest high tides (excludes wind driven
tides) with the sea level rise amount. These data should be used only as a
screening-level tool for management decisions. As with all remotely sensed data,
all features should be verified with a site visit. The dataset is provided "as
is," without warranty to its performance, merchantable state, or fitness for any
particular purpose. The entire risk associated with the results and performance
of this dataset is assumed by the user. This dataset should be used strictly as
a planning reference and not for navigation, permitting, or other legal
purposes.

```{r setup}

library(tidyverse)
library(sf)

path <- "/home/ajackson/Dropbox/Rprojects/TexasDioceseCreationCare/Data/"
output_path <- "/home/ajackson/Dropbox/projects/CeationCareTaskforce/Evacuation/"

Parishes <- readRDS(paste0(path, "Parishes_plus_Census_tracts.rds"))

Census <- readRDS(paste0(path, "Census_data_for_Texas_by_tract.rds"))

```

##    Read in sea level files

```{r read}

layers <- sf::st_layers(dsn = paste0(path, "sea_level/TX_North2_slr_final_dist.gdb"))

# foo1 <- sf::st_read(dsn = paste0(path, "sea_level/TX_North2_slr_final_dist.gdb"),
#                                  layer = "TX_North2_low_5ft")
# foo1 <- sf::st_make_valid(foo1)
# 
# foo1b <- rmapshaper::ms_simplify(foo1)
# foo1b <- sf::st_make_valid(foo1b)
# 
# foo2 <- sf::st_read(dsn = paste0(path, "sea_level/TX_North2_slr_final_dist.gdb"),
#                                  layer = "TX_North2_slr_5ft")
# foo2 <- sf::st_make_valid(foo2)
# foo2b <- rmapshaper::ms_simplify(foo2)
# foo2b <- sf::st_make_valid(foo2b)

OneFoot <- sf::st_read(dsn = paste0(path, "sea_level/TX_North2_slr_final_dist.gdb"),
                                 layer = "TX_North2_slr_1ft")
OneFoot <- sf::st_make_valid(OneFoot)
OneFoot <- rmapshaper::ms_simplify(OneFoot)
OneFoot <- sf::st_make_valid(OneFoot)
saveRDS(OneFoot, paste0(path, "sea_level/OneFoot_Tx_North2.rds"))

TwoFoot <- sf::st_read(dsn = paste0(path, "sea_level/TX_North2_slr_final_dist.gdb"),
                                 layer = "TX_North2_slr_2ft")
TwoFoot <- sf::st_make_valid(TwoFoot)
TwoFoot <- rmapshaper::ms_simplify(TwoFoot)
TwoFoot <- sf::st_make_valid(TwoFoot)
saveRDS(TwoFoot, paste0(path, "sea_level/TwoFeet_Tx_North2.rds"))

FiveFoot <- sf::st_read(dsn = paste0(path, "sea_level/TX_North2_slr_final_dist.gdb"),
                                 layer = "TX_North2_slr_5ft")
FiveFoot <- sf::st_make_valid(FiveFoot)
FiveFoot <- rmapshaper::ms_simplify(FiveFoot)
FiveFoot <- sf::st_make_valid(FiveFoot)
saveRDS(FiveFoot, paste0(path, "sea_level/FiveFeet_Tx_North2.rds"))

TenFoot <- sf::st_read(dsn = paste0(path, "sea_level/TX_North2_slr_final_dist.gdb"),
                                 layer = "TX_North2_slr_10ft")
TenFoot <- sf::st_make_valid(TenFoot)
TenFoot <- rmapshaper::ms_simplify(TenFoot)
TenFoot <- sf::st_make_valid(TenFoot)
saveRDS(TenFoot, paste0(path, "sea_level/TenFeet_Tx_North2.rds"))


#   make a diagnostic map

tmap::tmap_options(basemaps="OpenStreetMap")

tmap::tmap_mode("view") # set mode to interactive plots

tmap::tm_shape(TwoFoot) + 
  tmap::tm_fill(col="red", title = "2 ft slr", alpha=0.3, style="pretty")+
  tmap::tm_borders(lwd=0.5) +
tmap::tm_shape(OneFoot) + 
  tmap::tm_fill(col="green", title = "1 ft slr", alpha=0.3, style="pretty")+
  tmap::tm_borders(lwd=0.5) 
  




```












