---
title: "Smoke"
author: "Alan Jackson"
format: html
editor: source
---

## Look at smoke risks

Smoke data from https://www.stanfordecholab.com/wildfire_smoke

Paper is:
Daily Local-Level Estimates of Ambient Wildfire Smoke PM2.5 for the Contiguous US

by Marissa L. Childs*, Jessica Li, Jeffrey Wen, Sam Heft-Neal, Anne Driscoll, 
Sherrie Wang, Carlos F. Gould, Minghao Qiu, Jennifer Burney, and Marshall Burke

https://pubs.acs.org/doi/abs/10.1021/acs.est.2c02934

2006-2020 is period covered

```{r setup}

library(tidyverse)
library(sf)
library(duckdb)

Arc_CRS <- "EPSG:4735"
google_CRS <- "EPSG:4326"

path <- "/home/ajackson/Dropbox/Rprojects/TexasDioceseCreationCare/Data/"
Curated_path <- "/home/ajackson/Dropbox/Rprojects/Curated_Data_Files/Texas_Counties/"
out_path <- "/home/ajackson/Dropbox/projects/CeationCareTaskforce/Maps_and_tables/Smoke/"

Convocations <- readRDS(paste0(path,"Convocations.rds"))  

Parishes <- readRDS(paste0(path, "Parishes.rds"))

Smoke <- readRDS(paste0(path, "Smoke/smokePM2pt5_predictions_daily_county_20060101-20201231.rds"))

Counties <- readRDS(paste0(Curated_path, "County_Outlines.rds"))

```

##    Trim data to just Texas and summarize by county

Also do days by county by year

```{r}

Smoke <- inner_join(as_tibble(Counties %>% sf::st_drop_geometry()), 
                    Smoke, by=c('FIPS_ST_CN'='GEOID'))   

#   use duckdb to speed things up

con <- DBI::dbConnect(duckdb::duckdb(), dbdir = "duckdb")
duckdb::dbWriteTable(con, "Smoke", Smoke)
Smoke_db <- tbl(con, "Smoke")

Smoke_days_db <- Smoke_db %>% 
  group_by(CNTY_NM) %>% 
    summarize(Smoke_days=n())

Smoke_days <- collect(Smoke_days_db) %>% 
  left_join(., Counties, by="CNTY_NM") %>%  
  st_as_sf()

```

##    Make a map

```{r}

tmap::tmap_options(basemaps="OpenStreetMap")

tmap::tmap_mode("view") # set mode to interactive plots

tmap::tm_shape(Smoke_days) + 
  tmap::tm_fill(col = "Smoke_days", title = "Days with smoke, 2006-2020", alpha=0.5, style="pretty")+
  tmap::tm_borders(lwd=0.2) +
  tmap::tm_shape(Parishes) + 
  tmap::tm_dots()
```

##        Static map

```{r}

box <-sf::st_bbox(terrainr::add_bbox_buffer(Convocations, distance=5000)) # 5 kilometers

Base_basemapR <- basemapR::base_map(box, basemap="mapnik", increase_zoom=2)

Smoke_days %>% 
  mutate(Smoke_days=Smoke_days/17) %>% 
  sf::st_crop(box) %>% 
  ggplot() +
  Base_basemapR +
  geom_sf(aes(fill=Smoke_days) ,alpha=0.5) +
  scale_fill_gradient(name="Avg Days/Yr with Smoke", 
                      na.value = "transparent", 
                      low="yellow",
                      high="brown",
                      limits=c(300/17, 1300/17)) +
  geom_sf(data=Parishes, show.legend=FALSE) +
  # ggsflabel::geom_sf_label_repel(data=Convo_parishes, aes(label=Label), 
  #                                size=2, max.overlaps=25, force=80) +
  labs(title="Average Smoke Days per Year",
       subtitle=paste("from 2006 to 2020 by county")) +
  annotate("text", label="
         Data sourced from https://www.stanfordecholab.com/wildfire_smoke 
         Map a product of Alan Jackson", 
         x=-Inf, 
         y=-Inf,
         hjust="inward", vjust="inward", size = 2)

  ggsave(paste0(out_path, "Smoke_history.jpg"))
  
```







