---
title: "Smoke"
author: "Alan Jackson"
format: html
editor: source
---

## Look at smoke risks

Smoke data from https://www.stanfordecholab.com/wildfire_smoke

Paper is:
Daily Local-Level Estimates of Ambient Wildfire Smoke PM2.5 for the Contiguous US

by Marissa L. Childs*, Jessica Li, Jeffrey Wen, Sam Heft-Neal, Anne Driscoll, 
Sherrie Wang, Carlos F. Gould, Minghao Qiu, Jennifer Burney, and Marshall Burke

https://pubs.acs.org/doi/abs/10.1021/acs.est.2c02934

2006-2020 is period covered

```{r setup}

library(tidyverse)
library(sf)
library(duckdb)

Arc_CRS <- "EPSG:4735"
google_CRS <- "EPSG:4326"

path <- "/home/ajackson/Dropbox/Rprojects/TexasDioceseCreationCare/Data/"
Curated_path <- "/home/ajackson/Dropbox/Rprojects/Curated_Data_Files/Texas_Counties/"
out_path <- "/home/ajackson/Dropbox/projects/CeationCareTaskforce/Maps_and_tables/Smoke/"

Convocations <- readRDS(paste0(path,"Convocations.rds"))  

Parishes <- readRDS(paste0(path, "Parishes.rds"))

Smoke <- readRDS(paste0(path, "Smoke/smokePM2pt5_predictions_daily_county_20060101-20201231.rds"))

Counties <- readRDS(paste0(Curated_path, "County_Outlines.rds"))

```

##    Trim data to just Texas and summarize by county

Also do days by county by year

```{r}

Smoke <- inner_join(as_tibble(Counties %>% sf::st_drop_geometry()), 
                    Smoke, by=c('FIPS_ST_CN'='GEOID'))   

#   use duckdb to speed things up

con <- DBI::dbConnect(duckdb::duckdb(), dbdir = "duckdb")
duckdb::dbWriteTable(con, "Smoke", Smoke)
Smoke_db <- tbl(con, "Smoke")

Smoke_days_db <- Smoke_db %>% 
  group_by(CNTY_NM) %>% 
    summarize(Smoke_days=n())

Smoke_days <- collect(Smoke_days_db) %>% 
  left_join(., Counties, by="CNTY_NM") %>%  
  st_as_sf()

```

##    Make a map

```{r}

tmap::tmap_options(basemaps="OpenStreetMap")

tmap::tmap_mode("view") # set mode to interactive plots

tmap::tm_shape(Smoke_days) + 
  tmap::tm_fill(col = "Smoke_days", title = "Days with smoke, 2006-2020", alpha=0.5, style="pretty")+
  tmap::tm_borders(lwd=0.2) +
  tmap::tm_shape(Parishes) + 
  tmap::tm_dots()
```

##        Static map

```{r}

box <-sf::st_bbox(terrainr::add_bbox_buffer(Convocations, distance=5000)) # 5 kilometers

Base_basemapR <- basemapR::base_map(box, basemap="mapnik", increase_zoom=2)

Smoke_days %>% 
  mutate(Smoke_days=Smoke_days/17) %>% 
  sf::st_crop(box) %>% 
  ggplot() +
  Base_basemapR +
  geom_sf(aes(fill=Smoke_days) ,alpha=0.5) +
  scale_fill_gradient(name="Avg Days/Yr with Smoke", 
                      na.value = "transparent", 
                      low="yellow",
                      high="brown",
                      limits=c(300/17, 1300/17)) +
  geom_sf(data=Parishes, show.legend=FALSE) +
  # ggsflabel::geom_sf_label_repel(data=Convo_parishes, aes(label=Label), 
  #                                size=2, max.overlaps=25, force=80) +
  labs(title="Average Smoke Days per Year",
       subtitle=paste("from 2006 to 2020 by county")) +
  annotate("text", label="
         Data sourced from https://www.stanfordecholab.com/wildfire_smoke 
         Map a product of Alan Jackson", 
         x=-Inf, 
         y=-Inf,
         hjust="inward", vjust="inward", size = 2)

  ggsave(paste0(out_path, "Smoke_history.jpg"))
  
```

##        Table of worst parishes

```{r}

All_Parishes_Smoke <- 
inner_join(Parishes %>% st_drop_geometry(), 
           Smoke_days %>% st_drop_geometry(), 
           by=c("County"="CNTY_NM")) %>%
  select(Label, City, Convocation, Smoke_days) %>% 
  mutate(Smoke_days=round(Smoke_days/17, 0)) %>% 
  arrange(-Smoke_days)

# Histogram of above

All_Parishes_Smoke %>% 
  ggplot(aes(x=Smoke_days)) +
  geom_histogram()

table <- 
inner_join(Parishes %>% st_drop_geometry(), 
           Smoke_days %>% st_drop_geometry(), 
           by=c("County"="CNTY_NM")) %>%
  select(Label, City, Convocation, Smoke_days) %>% 
  mutate(Smoke_days=round(Smoke_days/17, 0)) %>% 
  filter(Smoke_days>60) %>% 
  arrange(-Smoke_days) %>% 
  gt::gt() %>% 
  gt::cols_label(
    Label="Church",
    City="City",
    Convocation="Convocation",
    Smoke_days="Smoke Days per Year, Top Parishes"
  )

gt::gtsave(table, paste0(out_path, "Smoke_days.docx"))
```


##        Plot up what time of year we see smoke - a monthly bar chart

```{r}

Smoke_month_db <- Smoke_db %>% 
  mutate(Month=lubridate::month(date)) %>% 
  group_by(CNTY_NM, Month) %>% 
    summarize(Smoke_days=n()/17)

Smoke_Months <- collect(Smoke_month_db) %>% 
  inner_join(., Convocations, by=c("CNTY_NM"="County")) %>%  
  sf::st_drop_geometry() 

Smoke_Months %>% 
  group_by(Month) %>% 
    summarise(Smoke_days_avg=mean(Smoke_days),
              Smoke_days_sd=2*sd(Smoke_days)) %>% 
  ggplot(aes(x=Month, y=Smoke_days_avg)) +
    geom_errorbar(aes(ymin=pmax(0, (Smoke_days_avg-Smoke_days_sd)), 
                      ymax=Smoke_days_avg+Smoke_days_sd))+
    geom_point() +
    geom_line() +
    scale_x_continuous(
      breaks = seq_along(month.name), 
      labels = month.abb
    ) +
    labs(title="Average Texas Smoke Days per Month per Year, and Range",
         y="Smoke Days per Year") +
  annotate("text", label="
         Data sourced from https://www.stanfordecholab.com/wildfire_smoke 
         Map a product of Alan Jackson", 
         x=-Inf, 
         y=-Inf,
         hjust="inward", vjust="inward", size = 2)

ggsave(paste0(out_path, "Smoke_days_per_month.jpg"))
```





