---
title: "Wildfire_grids.qmd"
author: "Alan Jackson"
format: html
editor: source
---

## Read in the fire potential grids and analyze

```{r setup}

library(tidyverse)

Arc_CRS <- "EPSG:4735"
google_CRS <- "EPSG:4326"

path <- "/home/ajackson/Dropbox/Rprojects/TexasDioceseCreationCare/Data/"
#Curated_path <- "/home/ajackson/Dropbox/Rprojects/Curated_Data_Files/Texas_Counties/"
out_path <- "/home/ajackson/Dropbox/projects/CeationCareTaskforce/Maps_and_tables/Fire/"

Convocations <- readRDS(paste0(path,"Convocations.rds"))  

Parishes <- readRDS(paste0(path, "Parishes.rds"))# %>% 
  # mutate(City=stringr::str_trim(City) ) %>% 
  # mutate(Label=stringr::str_trim(Label) )

```

### convert gdb to geotiff

Using code from https://github.com/r-barnes/ArcRasterRescue to extract the raster 
from the gdb file as a geotiff

Syntax A: arc_raster_rescue.exe <File Geodatabase> 

Syntax B: arc_raster_rescue.exe <File Geodatabase> <Raster> <Output Name>

Syntax A will list all of the rasters in the data set along with selection numbers for use with Syntax B.

Syntax B, given an FGDB and raster selection number, will output the raster to the indicated output file.

Syntax B also accepts raster names, as listed by Syntax A, as inputs for <Raster>

NOTE: The geodatabase path must end with a slash!

EXAMPLE: ./arc_raster_rescue.exe path/to/geodatabase.gdb/ dem03 /z/out.tif

Select layer 6 Threat_UP_i

###    Read in geotiffs

Wildfire Threat
Wildfire Threat is the likelihood of a wildfire occurring or burning into an area.

Threat is derived by combining a number of landscape characteristics including surface and canopy fuels, resultant fire behavior, historical fire occurrence, percentile weather derived from historical weather observations, and terrain conditions. These inputs are combined using analysis techniques based on established fire science.

The measure of wildfire threat used in the Texas Wildfire Risk Assessment (TWRA) is based on the Wildland Fire Susceptibility Index (WFSI). WFSI combines the probability of an acre igniting (Wildfire Ignition Density), and the expected final fire size based on rate of spread in four percentile weather categories. WFSI is defined as the likelihood of an acre burning.

Since all areas in Texas have WFSI calculated consistently, it allows for comparison and ordination of areas across the entire state. For example, a high threat area in East Texas is equivalent to a high threat area in West Texas.

The threat map is derived at a 30-meter resolution. This scale of data was chosen to be consistent with the accuracy of the primary surface fuels dataset used in the assessment. While not appropriate for site specific analysis, it is ideal for state and regional scale analysis.

```{r}

files <- list.files(paste0(path, "Fire/" ),
                    pattern=".*geotiff")

fire <- terra::aggregate(terra::rast(paste0(path, "Fire/", files[1])), 
                         fact=10, fun="max", na.rm=TRUE)
# foo <- terra::aggregate(terra::rast(paste0(path, "Fire/", files[2])),
#                          fact=10, fun="max", na.rm=TRUE)

for (i in 2:length(files)-1){
  print(paste("---", i, "---"))
  foo <- terra::aggregate(terra::rast(paste0(path, "Fire/", files[i])),
                         fact=10, fun="max", na.rm=TRUE)
  fire <- terra::merge(fire, foo)
}
foo <- terra::aggregate(terra::rast(paste0(path, "Fire/", files[14])),
                         fact=10, fun="max", na.rm=TRUE)
foo <- terra::resample(foo, fire, method="near")
fire <- terra::merge(fire, foo)

tmap::tmap_options(basemaps="OpenStreetMap")

tmap::tmap_mode("view") # set mode to interactive plots

tmap::tm_shape(fire) + 
    #tmap::tm_sf(col="ConvoColor", alpha=0.3) +
  # tmap::tm_fill(title = "Threat", alpha=0.3, style="pretty") +
  tmap::tm_raster(alpha=0.3,
                  title = "Wildfire Threat",
                  palette = "-RdYlGn") +
  tmap::tm_shape(Parishes)+
  tmap::tm_dots()
  # tmap::tm_borders(lwd=3, col="red") 
```

###   Let's use fire likelihood as a proxy for smoke

```{r}

#   Turn danger likelihood into a factor

# fire2 <- fire %>% 
# mutate(Risk=factor(Risk, levels=c('minimal', 'very low', 'very low/low',
#                                             'low', 'low/moderate',
#                                             'moderate', 'moderate/high', 'high',
#                                             'high/very high'))) %>% 
#   sf::st_as_sf()
fire2 <- fire
fire2 <- terra::focal(fire2, window=21, fun=max, na.policy="only", na.rm=TRUE)
fire2 <- fire2 - 1
levels(fire2) <- c('low', 'low/moderate',
                   'moderate', 'moderate/high', 'high',
                   'high/very high', 'very high')


#   Make a ggplot 

fire2 <- terra::project(fire2, google_CRS)
bbox <- sf::st_bbox(fire2)

Base_basemapR <- basemapR::base_map(bbox, basemap="mapnik", increase_zoom=2)

#   All points

Parishes %>% 
  ggplot() +
  Base_basemapR +
  # geom_sf(color="black", shape=1 ) +
  geom_sf(color="black") +
  tidyterra::geom_spatraster(data=fire2, alpha=0.3) +
  # scale_color_gradientn(colors=RColorBrewer::brewer.pal(7, "RdOrYl"),
  scale_fill_brewer(palette="RdYlGn", direction=-1,
  # scale_color_gradientn(colors=scales::brewer_pal(type="seq", palette="RdOrYl",
  #                       direction=+1),
                           name="Fire Likelihood",
                           na.value = "grey100") +
    labs(title=paste("Fire Likelihood for Churches in the Diocese"),
         x="Longitude",
       y="Latitude") +
  annotate("text", label=
           "Data sourced from Texas A&M Forest Service 
           Wildfire Risk Explorer
           Map a product of Alan Jackson", 
           x=-Inf, 
           y=-Inf,
           hjust="inward", vjust="inward", size = 2) +
  coord_sf(xlim=c(bbox$xmin, bbox$xmax),c(bbox$ymin, bbox$ymax))

  ggsave(paste0(out_path, "Fire_likelihood_All_Churches.jpg"))
```





