---
title: "Wildfire_grids.qmd"
author: "Alan Jackson"
format: html
editor: source
---

## Read in the fire potential grids and analyze

```{r setup}

library(tidyverse)

Arc_CRS <- "EPSG:4735"

path <- "/home/ajackson/Dropbox/Rprojects/TexasDioceseCreationCare/Data/"
Curated_path <- "/home/ajackson/Dropbox/Rprojects/Curated_Data_Files/Texas_Counties/"
out_path <- "/home/ajackson/Dropbox/projects/CeationCareTaskforce/Maps_and_tables/Fire/"

Convocations <- readRDS(paste0(path,"Convocations.rds"))  

Parishes <- readRDS(paste0(path, "Parishes.rds")) %>% 
  mutate(City=stringr::str_trim(City) ) %>% 
  mutate(Label=stringr::str_trim(Label) )

```

### convert gdb to geotiff

Using code from https://github.com/r-barnes/ArcRasterRescue to extract the raster 
from the gdb file as a geotiff

Syntax A: arc_raster_rescue.exe <File Geodatabase> 

Syntax B: arc_raster_rescue.exe <File Geodatabase> <Raster> <Output Name>

Syntax A will list all of the rasters in the data set along with selection numbers for use with Syntax B.

Syntax B, given an FGDB and raster selection number, will output the raster to the indicated output file.

Syntax B also accepts raster names, as listed by Syntax A, as inputs for <Raster>


NOTE: The geodatabase path must end with a slash!

EXAMPLE: ./arc_raster_rescue.exe path/to/geodatabase.gdb/ dem03 /z/out.tif

###    Read in geotiffs

```{r}
fire <- terra::rast(paste0(path, "Fire/Threat_UP.geotiff"))
```





